Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-08-06 submission/manuscript.Rmd
checking if new vsc works
update kdm5c-bound promoters results

The majority of promoters bound by KDM5C in PNCs were also bound in EpiLCs (513 shared promoters), however a large portion of gene promoters were only bound by KDM5C in EpiLCs (3677 EpiLC only promoters) (Figure 4B). We then performed gene ontology analysis to compare the known functions of genes with KDM5C bound to their promoter in EpiLCs and PNCs. While there were no ontologies significantly enriched for genes only bound by KDM5C in PNCs, gene ontologies for genes shared between PNCs and EpiLCs were enriched for functions involving nucleic acid turnover, such as deoxyribonucleotide metabolic process (GO:0009262, p.adjust = `r signif(getGO("deoxyribonucleotide metabolic process", GO_KDM5C_ChIP, 'p.adjust', "Shared"), 3)`) (Figure 4C). Germline-specific ontologies were only enriched in promoters unique to EpiLCs, such as meiotic nuclear division (GO: 0007127 p.adjust = `r signif(getGO("meiosis I", GO_KDM5C_ChIP, 'p.adjust', "EpiLC only"), 3)`) and meiotic cell cycle process (GO:1903046, p.adjust = `r signif(getGO("meiotic cell cycle process", GO_KDM5C_ChIP, 'p.adjust', "EpiLC only"), 3)`) (Figure 3C). We then evaluated KDM5C binding around the transcription start site (TSS) of all germline-enriched genes. While KDM5C was bound to a subset of germline gene promoters in EpiLCs, it was not bound to any in PNCs (Figure 4D-E). Together, these results demonstrate KDM5C is significantly enriched at a subset of germline gene promoters in EpiLCs, including meiotic genes, but does not directly regulate germline genes in neurons.

@@

## KDM5C promotes _de novo_ DNA methylation at germline genes

>>>>

The majority of promoters bound by KDM5C in PNCs were also bound in EpiLCs (513 shared promoters), however a large portion of gene promoters were only bound by KDM5C in EpiLCs (3677 EpiLC only promoters) (Figure 4B). We then performed gene ontology analysis to compare the known functions of genes with KDM5C bound to their promoter in EpiLCs and PNCs. While there were no ontologies significantly enriched for genes only bound by KDM5C in PNCs, gene ontologies for genes shared between PNCs and EpiLCs were enriched for functions involving nucleic acid turnover, such as deoxyribonucleotide metabolic process (GO:0009262, p.adjust = `r signif(getGO("deoxyribonucleotide metabolic process", GO_KDM5C_ChIP, 'p.adjust', "Shared"), 3)`) (Figure 4C). Germline-specific ontologies were only enriched in promoters unique to EpiLCs, such as meiotic nuclear division (GO: 0007127 p.adjust = `r signif(getGO("meiosis I", GO_KDM5C_ChIP, 'p.adjust', "EpiLC only"), 3)`) and meiotic cell cycle process (GO:1903046, p.adjust = `r signif(getGO("meiotic cell cycle process", GO_KDM5C_ChIP, 'p.adjust', "EpiLC only"), 3)`) (Figure 3C). We then evaluated KDM5C binding around the transcription start site (TSS) of all germline-enriched genes. In EpiLCs, We observed modest KDM5C signal at about half of all germline genes (Figure 4D). Based on our ChIP-seq peak criteria, KDM5C was strongly bound to about 10% of germline gene promoters in EpiLCs (Figure 4E). In condordance with our gene ontology results, we did not observe KDM5C accumiulation at any germline gene promtoers in PNCs (Figure 4D). Together, these results demonstrate KDM5C is significantly enriched at a subset of germline gene promoters in EpiLCs, including meiotic genes, but does not directly regulate germline genes in neurons.

- read in the homer results txt file above here

- Germline genes repressed by other histone modifying enzymes have shared genomic features around the TSS to suppress their expression, such as motifs for E2F6/DP1 (E2F) and MGA/MAX (E-box) binding.
	- We asessed genomic features enriched at KDM5C-bound versus unbound gemrline gene promoters using HOMER.
		- window I analyzed

@@

## KDM5C promotes _de novo_ DNA methylation at germline genes
- if there are differences, say at germline gene CpG islands
